---
title: "askMe Data User Guide"
shorttitle: "sesameData guide"
package: sesameData
output: rmarkdown::html_vignette
fig_width: 6
fig_height: 5
vignette: >
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteIndexEntry{sesameData User Guide}
    %\VignetteEncoding{UTF-8}
---

# Introduction

`sesameData` package provides associated data for sesame package. This includes
example data for testing and instructional purpose, as we ll as probe annotation
for different Infinium platforms.

```{r, message=FALSE, warning=FALSE}
library(sesameData)
library(GenomicRanges)
```

# Data from ExperimentHub

Titles of all the available data can be shown with:
```{r}
head(sesameDataList())
```

## Local caching

Each sesame datum from ExperimentHub is accessible through the `sesameDataGet`
interface. It should be noted that all data must be pre-cached to local disk
before they can be used.  This design is to prevent conflict in annotation data
caching and remove internet dependency. Caching needs only be done once per
sesame/sesameData installation. One can cache data using

```{r, include = FALSE}
sesameDataCacheExample()        # examples
```

```{r}
sesameDataCache()
```

Once a data object is loaded, it is stored to a tempoary cache, so that the data
doesn't need to be retrieved again next time we call `sesameDataGet`. This
design is meant to speeed up the run time.

For example, the annotation for HM27 can be retrieved with the title: 
```{r}
HM27.address <- sesameDataGet('HM27.address')
```

## In-memory caching

It's worth noting that once a data is retrieved through the `sesameDataGet`
inferface (below), it will stay in memory so next time the object will be
returned immediately. This design avoids repeated disk/web retrieval. In some
rare situation, one may want to redo the download/disk IO, or empty the cache
to save memory. This can be done with:

```{r}
sesameDataGet_resetEnv()
```

## Transcript models

Sesame provides some utility functions to process transcript models, which can
be represented as data.frame, GRanges and GRangesList objects. For example,
`sesameData_getTxnGRanges` calls `sesameDataGet("genomeInfo.mm10")$txns` to
retrieve a transcript-centric GRangesList object from GENCODE including its
gene annotation, exon and cds (for protein-coding genes). It is then turned
into a simple GRanges object of transcript:

```{r}
txns_gr <- sesameData_getTxnGRanges("mm10")
txns_gr
```

The returned GRanges object does not contain the exon coordinates. We can
further collapse different transcripts of the same gene (isoforms) to gene
level. Gene start is the minimum of all isoform starts and end is the maximum
of all isoform ends.

```{r}
genes_gr <- sesameData_txnToGeneGRanges(txns_gr)
genes_gr
```

# Annotate probes

One can annotate given probe ID using any genomic features stored in GRanges
objects. For example, the following demonstrate the annotation of 500 random
Mammal40 probes for gene promoters.

```{r}
probes <- names(sesameData_getManifestGRanges("Mammal40"))[1:500]
head(probes) # our input
txns <- sesameData_getTxnGRanges("hg38")
pm <- promoters(txns, upstream = 1500, downstream = 1500)
pm <- pm[pm$transcript_type == "protein_coding"]
sesameData_annoProbes(probes, pm, column = "gene_name")
```

# Manifest from Github

Sesame provides access to array manfiest stored as GRanges object. These
GRanges object are converted from the raw tsv files on our [array annotation
website](http://zwdzwd.github.io/InfiniumAnnotation) using the [conversion
code](https://tinyurl.com/8bt9dssc)

```{r}
gr <- sesameData_getManifestGRanges("HM450")
length(gr)
```

Note that by default the GRanges object exclude decoy sequence probes (e.g.,
_alt, and _random contigs). To include them, we need to use the `decoy = TRUE`
option in sesameData_getManifestDF.

# Subset probes

One can directly get probes from different parts of the genome.

```{r message = FALSE, warning = FALSE}
library(GenomicRanges)
```

```{r}
regs <- GRanges('chr5', IRanges(135313937, 135419936))
sesameData_getProbesByRegion(regs, platform = 'Mammal40')
sesameData_getProbesByChromosome('chrX', platform = 'Mammal40')
sesameData_getAutosomeProbes("Mammal40")
sesameData_getProbesByGene('DNMT3A', "Mammal40", upstream=500)
sesameData_getProbesByTSS('DNMT3A', "Mammal40")
```

One can also get all TSS probes by

```{r}
TSSprobes = sesameData_getProbesByTSS(NULL, "Mammal40")
```

# Get nearby genes

```{r}
sesameData_getGenesByProbes(c("cg14620903","cg22464003"), max_distance = 10000)
```

# Installation

From Bioconductor
```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("sesameData")
```

Development version can be installed from github
```{r, eval=FALSE}
BiocManager::install("zwdzwd/sesameData")
```

```{r}
sessionInfo()
```
